Constraint Programming Algorithms used in Gecode

### 2014-06-17 paper.tex
fixed title of seminar
improvements

\begin{document}
	\title{Constraint Programming Algorithms used in Gecode}
	\author{Benedikt~Schmidt}
	\markboth{Advanced Seminar for Security in Information Technology, Summer Term 2014}%
	{Benedikt Schmidt: Constraint Programming - Inside \emph{Gecode}}	
	\maketitle

@@

	\begin{figure}[b]
		\center
		\begin{tikzpicture}
			\draw (0, 0) rectangle (1, 1);
			\node at (0, -0.3) {($x_1$, $y_1$)};
			\draw[arrows=<->] (1.2, 0) -- (1.2, 1);
			\node at (1.5, 0.5) {$d_1$};
			\draw (1.5, 1.5) rectangle(2, 2);
			\node at (1.5, 1.2) {($x_2$, $y_2$)};
			\draw[arrows=<->] (2.2, 1.5) -- (2.2, 2);
			\node at (2.5, 1.75) {$d_2$};
		\end{tikzpicture}
		\caption{Two squares which should not overlap}
		\label{fig:squares}
	\end{figure}

@@

	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: variables, constraints, objective functions, search trees, constraint propagation and search algorithms. As example of an implementation I will refer to \emph{Gecode} \cite{gecode}.

	\section{State of the Art}
	\subsection{Variable Domains}
	In constraint programming a variable is connected to a set of possible values, the so called domain. Depending on the domain, a variable is called an integer, Boolean or floating variable. Possible ranges for a variable $x$ of each domain are:
	\begin{itemize}
		\item Integer: finite or infinite, e.g. $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$
		\item Bool: $x \in \{\text{true}, \text{false}\}$
		\item Float: intervals or unions of intervals, e.g. $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$
	\end{itemize}
	This distinction is fundamental as especially for floating variables a combinatorial approach to constraint programming is impossible.

	\subsection{Constraint Types}
	The choice of constraint types and how comfortable it is to specify them varies from the programming language in which the problem is implemented. For sure, there is also a difference in the interfaces of the libraries for constraint programming. Typical constraints are
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}

@@

	\begin{thebibliography}{1}
		\bibitem{handbookCP}
		F.~Rossi, P.~van~Beek and T.~Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN-13: 978-0-444-52726-4
		\bibitem{allDifferent}
		W.~van~Hoeve, \emph{The Alldifferent Constraint: A Survey}, http://www.andrew.cmu.edu/user/vanhoeve/papers/alldiff.pdf
		\bibitem{trendsInCP}
		F.~Benhamou, N.~Jussien, B.~A.~O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN-13: 978-1-905209-97-2
		\bibitem{gecode}
		http://www.gecode.org/
		\bibitem{programmingGecode}
		C.~Schulte, G.~Tack and M.~Z.~Lagerkvist, \emph{Modeling and Programming with Gecode}, http://www.gecode.org/doc-latest/MPG.pdf
		\bibitem{limitedDiscrepancy}
		W.~D.~Harvey, M.~L.~Ginsberg, \emph{Limited Discrepancy Search}, Morgan Kaufmann, 1995
		\bibitem{linearProgramming}
		L.~Cooper and D.~Steinberg, \emph{Methods and Applications of Linear Programming}, W.~B.~Saunder~Company, 1974, ISBN-10: 0-7216-2694-7
		\bibitem{artOfComputerProgramming}
		D.~E.~Knuth, \emph{The Art of Computer Programming: Volume 3, Sorting and Searching}, second edition, Addison-Wesley, 1998, ISBN-10: 0201896850, ISBN-13: 9780201896855
		\bibitem{designPatterns}
		E.~Gamma, R.~Helm, R.~Johnson, J.~Vlissides, \emph{Design Patterns. Elements of Reusable Object-Oriented Software}, first edition, Prentice Hall, 1994, ISBN-10: 0201633612
	\end{thebibliography}
\end{document}

>>>>

\begin{document}
	\title{Constraint Programming Algorithms used in Gecode}
	\author{Benedikt~Schmidt <benediktibk@aon.at>}
	\markboth{Advanced Seminar for VLSI Design Methods, Summer Term 2014}%
	{Benedikt Schmidt: Constraint Programming - Inside \emph{Gecode}}
	\maketitle

@@

	\begin{figure}[h]
		\center
		\begin{tikzpicture}
			\draw (0, 0) rectangle (1, 1);
			\node at (0, -0.3) {($x_1$, $y_1$)};
			\draw[arrows=<->] (1.2, 0) -- (1.2, 1);
			\node at (1.5, 0.5) {$d_1$};
			\draw (1.5, 1.5) rectangle(2, 2);
			\node at (1.5, 1.2) {($x_2$, $y_2$)};
			\draw[arrows=<->] (2.2, 1.5) -- (2.2, 2);
			\node at (2.5, 1.75) {$d_2$};
		\end{tikzpicture}
		\caption{Two squares which should not overlap}
		\label{fig:squares}
	\end{figure}

@@

	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: variables, constraints, objective functions, search trees, constraint propagation and search algorithms. As example of an implementation I will refer to \emph{Gecode}\footnote{http://www.gecode.org/}.

	\section{State of the Art}
	\subsection{Variable Domains}
	In constraint programming a variable is connected to a set of possible values, the so called domain. Depending on the domain, a variable is called an integer, Boolean or floating variable. Possible ranges for a variable $x$ of each domain are:
	\begin{itemize}
		\item Integer: finite or infinite, e.g. $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$
		\item Bool: $x \in \{\text{true}, \text{false}\}$
		\item Float: intervals or unions of intervals, e.g. $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$
	\end{itemize}
	This distinction is fundamental as especially for floating variables a combinatorial approach is impossible.

	\subsection{Constraint Types}
	The choice of constraint types and how comfortable it is to specify them varies from the programming language in which the problem is implemented. For sure, there is also a difference in the interfaces of the libraries for constraint programming. Typical constraints are:
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}

@@

	\begin{thebibliography}{1}
		\bibitem{handbookCP}
		F.~Rossi, P.~van~Beek and T.~Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN-13: 978-0-444-52726-4
		\bibitem{allDifferent}
		W.~van~Hoeve, \emph{The Alldifferent Constraint: A Survey}, http://www.andrew.cmu.edu/user/vanhoeve/papers/alldiff.pdf
		\bibitem{trendsInCP}
		F.~Benhamou, N.~Jussien, B.~A.~O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN-13: 978-1-905209-97-2
		\bibitem{programmingGecode}
		C.~Schulte, G.~Tack and M.~Z.~Lagerkvist, \emph{Modeling and Programming with Gecode}, http://www.gecode.org/doc-latest/MPG.pdf
		\bibitem{limitedDiscrepancy}
		W.~D.~Harvey, M.~L.~Ginsberg, \emph{Limited Discrepancy Search}, Morgan Kaufmann, 1995
		\bibitem{linearProgramming}
		L.~Cooper and D.~Steinberg, \emph{Methods and Applications of Linear Programming}, W.~B.~Saunder~Company, 1974, ISBN-10: 0-7216-2694-7
		\bibitem{artOfComputerProgramming}
		D.~E.~Knuth, \emph{The Art of Computer Programming: Volume 3, Sorting and Searching}, second edition, Addison-Wesley, 1998, ISBN-10: 0201896850, ISBN-13: 9780201896855
		\bibitem{designPatterns}
		E.~Gamma, R.~Helm, R.~Johnson, J.~Vlissides, \emph{Design Patterns. Elements of Reusable Object-Oriented Software}, first edition, Prentice Hall, 1994, ISBN-10: 0201633612
	\end{thebibliography}
\end{document}
